---
title: "jSDM"
author: "Aimara Planillo"
date: "2024-02-12"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# jSDM for mice microbiome    

We use 
* Response: the bacterial microbiome in the mice as response.
* Explanatory variables: environmental variables and mouse characteristics
* Spatial effects: longitude and latitude
* Random: year


### prepare workspace
```{r}
source("rscripts/source_pkgs.R") # r libraries

wrkdir <- getwd()
output_wd <- paste0(wrkdir, "/output/jSDM_20240212")
```

### load data
```{r}
# mice related data
mice_data <- read.csv("./intermediate_data/Table_mice_envcov_100m.csv")

# microbiome bacteria data
PS.TSS_filtered_jS <- readRDS("./intermediate_data/PSvar.rds")
PS.TSS_filtered_jS

sample_data(PS.TSS_filtered_jS@sam_data)
## filter data for bacteria
PS_bacteria <- subset_taxa(PS.TSS_filtered_jS, Kingdom %in%"Bacteria")

PS_bacteria@otu_table

## get data at family level
rank_names(PS_bacteria)[5]
PS_bacteria_fam <- tax_glom(PS_bacteria, taxrank=rank_names(PS_bacteria)[5])
PS_bacteria_fam

## 60 bacteria Families in 607 mice
```


## extract bacteria presence / absence data for each mouse
```{r get table with pa data}
## cahnge the row names of tax table and col names of otu tables for new ones so they are easier to merge
rownames(PS_bacteria_fam@tax_table) <- paste0("ASV", 1:nrow(PS_bacteria_fam@tax_table))
colnames(PS_bacteria_fam@otu_table) <- paste0("ASV", 1:ncol(PS_bacteria_fam@otu_table))

PS_bacteria_fam@otu_table
PS_bacteria_fam@tax_table

otut <- data.frame(PS_bacteria_fam@otu_table)
taxt <- data.frame(PS_bacteria_fam@tax_table)
taxt$Family

colnames(otut) <- taxt$Family

head(otut)

## make data presence / absence
otut_pa <- otut
otut_pa[otut_pa > 0] <- 1
 
head(otut_pa)

otut_pa <- otut_pa %>%
  tibble::rownames_to_column("mouseID")

```


```{r check duplicates}
## to remove duplicates, we need to transpose the data, merge the rows with the same name and transpose again to the original format
  
sampleID <- rownames(otut_pa)
taxaID <- colnames(otut_pa)

length(unique(taxaID)) # 60 Fam
length(unique(sampleID)) # 607 samples
# No dup
```

## prepare final mice data
Assign precipitation value to the year the mouse where captured
```{r}
head(mice_data)

mice_data %>% 
  group_by(jS_loc) %>% 
  # mutate(numId = seq(1:length(unique(jS_loc)))) 
  mutate(locID =  cur_group_id()) %>% 
  dplyr::select(jS_mouseID, locID)
 
 
## filter environmnetal variables and set pig density NA to 0
mice_data2 <- mice_data %>% 
  group_by(jS_loc) %>% 
  mutate(numlocID =  cur_group_id()) %>% 
  mutate(locID = paste0("loc_", numlocID)) %>% 
  ungroup() %>% 
  mutate(precipation_aug = case_when(
    jS_year == 2015 ~ prec_aug_2015,
    jS_year == 2016 ~ prec_aug_2016,
    jS_year == 2017 ~ prec_aug_2017,
    jS_year == 2018 ~ prec_aug_2018,
    TRUE ~NA)) %>% 
  mutate(jS_sex = as.factor(jS_sex), 
         municipality_code = as.factor(as.character(gem_char)), 
         clc_fct = as.factor(clc_fct)) %>% 
  mutate(pig_dens = replace_na(pig_dens, 0)) %>% 
  dplyr::select(-c(jS_loc, prec_aug_2015, prec_aug_2016, prec_aug_2017, prec_aug_2018, 
                   ID, CATO, PITO, POTO, gem_char))
mice_data2

## add bacteria
all_data <- mice_data2 %>% 
  left_join(otut_pa, by = c("jS_mouseID" = "mouseID"))

head(all_data)
```


# Run jSDM

For the JSDM we need:
- a matrix of species: sites in rows (mice) and species in columns (OTU)
- a data frame with the environmental covariates for sites: sites (mice) in rows, same order as above
- random structure
- spatial structure

## Set up variables and model structure

```{r}
colnames(all_data)

# response matrix 
my_response <- colnames(otut_pa)[-1] ## remove mouseID
response_pa <- all_data %>% 
  dplyr::select(all_of(my_response)) %>% 
  as.matrix()

# explanatory df
my_cov <- c("jS_sex", "jS_BMI", "jS_HI", "jS_Ntrapped", "imperv", "tcd", 
            "swf_perc", "dist_road", "dist_path", #"soil_moist", "soil_ph", 
            "cat_dens", "pig_dens", "pou_dens", "clc_fct", "precipation_aug", "municipality_code")
explanatory_var <- all_data %>% 
  dplyr::select(all_of(my_cov)) 

as.data.frame(explanatory_var[!complete.cases(explanatory_var),])

## spatial data (wihtout duplicates)
xyData <- all_data %>%
  group_by(locID) %>% 
  summarise(lon = unique(jS_lon), 
            lat = unique(jS_lat)) %>% 
  tibble::column_to_rownames("locID")

## two nested random effects
studyDesign <- data.frame(sample = as.factor(all_data$jS_mouseID), locality = as.factor(all_data$locID), 
                          year = as.factor(all_data$jS_year))


rL3 <- HmscRandomLevel(units = levels(all_data$jS_year)) # account for temporal correlation
rL2 <- HmscRandomLevel(sData = xyData, units = levels(all_data$locID)) # accounts for nested sampling
rL1 <- HmscRandomLevel(units = levels(all_data$jS_mouseID)) # obtain residual correlations

```


### Define MCMC parameters
```{r}
thin <- 10
samples <- 10000
transient <- 1000
nChains <- 3
verbose <- 1000

# Regression formula for environmental covariates

## there are many NA in soil ph and soil moisture is correlated to precipitation, thus we remove those variables

XFormula_OTU = ~ jS_sex + jS_BMI + jS_HI + jS_Ntrapped + imperv + tcd + swf_perc +
  swf_perc + dist_road + dist_path + #soil_ph + soil_moist +
  cat_dens + pig_dens + pou_dens + clc_fct + precipation_aug + municipality_code

```

## *BINOMIAL DISTRIBUTION* ~> PROBIT MODEL

### Fit models for PRESENCE/ABSENCE  data 

```{r}
m <- Hmsc(Y = response_pa, XData = explanatory_var, XFormula = XFormula_OTU,
           studyDesign=studyDesign, ranLevels=list("sample" = rL3,"locality"= rL2, "year" = rL3),
          distr = "probit")
```

## Run model
```{r}
# Run Markov Chains
# test
m <- sampleMcmc(m, thin = 1, samples = 50)

# the real model
m <- sampleMcmc(m, thin = thin, samples = samples, transient = transient, 
                nChains = nChains, verbose = verbose, nParallel = nChains)


model_wd <- paste0(getwd(), "/output/jSDM_bact_fam_pa_202402")
# Set a name for the model
filename <- paste0(model_wd, "/JSDM_mice_microbiome_bacteria_pa_20240214.rds")

#Save de model
saveRDS(m,file=filename) 
```