#### ------------------- ########
THIS IS FOR PLOTTING 
#### ------------------- ########

## Predicted responses 95CI

### Example full community
```{r}
m$XFormula
head(m$X)

Gradient.season <- constructGradient(m, focalVariable = "season")
predY <- predict(m, XData = Gradient.season$XDataNew, studyDesign = Gradient.season$studyDesignNew, 
                 ranLevels = Gradient.season$rLNew, expected = FALSE)

# Summed response to season
plotGradient(m, Gradient.season, pred=predY, measure="S", las=1,
             showData = TRUE, main='Species richness (measure="S")')

# Species 10
plotGradient(m, Gradient.season, pred=predY, measure="Y", index=3, las=1,
             showData = TRUE, main='Focal species occurrence (measure="Y", species = 3)')
# Community-weighted mean values of traits
# plotGradient(m, Gradient.season, pred=predY, measure="T", index=2, las=1,
#              showData = TRUE, main='Mean trait value (measure="T")')


# Summed response to tree
Gradient.tree <- constructGradient(m, focalVariable = "tree_cover_100")
predY.tree <- predict(m, XData = Gradient.tree$XDataNew, 
                 studyDesign = Gradient.tree$studyDesignNew, 
                 ranLevels = Gradient.tree$rLNew, expected = FALSE)
saveRDS(predY.tree, paste0(det_model_wd, "/pred_response_tree.rds"))

plotGradient(m, Gradient.tree, pred=predY.tree, measure="S", las=1,
             showData = TRUE, main='Total abundance')
```

-----------------------
More options
-----------------------


# 3. Species Associations plot - A

Hierarchical Edge Bundle, simple approach
```{r}
# Libraries
library(ggraph)
library(igraph)
library(tidyverse)
library(RColorBrewer)
 
# create a data frame giving the hierarchical structure of your individuals
d1 <- data.frame(from = "origin", to=paste("group", seq(1,10), sep=""))
d2 <- data.frame(from = rep(d1$to, each = 6), to = unique(associations$sp1))
edges <- rbind(d1, d2)
 
# create a dataframe with connection between leaves (individuals)
all_leaves <- unique(associations$sp1)
connect <- data.frame( from = associations$sp1, 
                       to = associations$sp2, 
                       value = associations$mean, 
                       support = associations$support)

connect <- connect %>% 
  filter(support < -0.95 | support > 0.95)
 
# create a vertices data.frame. One line per object of our hierarchy
vertices  <-  data.frame(
  name = unique(c(as.character(edges$from), as.character(edges$to))) , 
  value = runif(length(edges$from)+1)
) 
# Let's add a column with the group of each name. It will be useful later to color points
vertices$group  <-  edges$from[ match( vertices$name, edges$to ) ]



#Let's add information concerning the label we are going to add: angle, horizontal adjustement and potential flip
#calculate the ANGLE of the labels
vertices$id <- NA
myleaves <- which(is.na( match(vertices$name, edges$from) ))
nleaves <- length(myleaves)
vertices$id[ myleaves ] <- seq(1:nleaves)
vertices$angle <- 90 - 360 * vertices$id / nleaves
 
# calculate the alignment of labels: right or left
# If I am on the left part of the plot, my labels have currently an angle < -90
vertices$hjust <- ifelse( vertices$angle < -90, 1, 0)
 
# flip angle BY to make them readable
vertices$angle <- ifelse(vertices$angle < -90, vertices$angle+180, vertices$angle)



# Create a graph object
mygraph <- igraph::graph_from_data_frame( edges, vertices=vertices )
 
# The connection object must refer to the ids of the leaves:
from  <-  match( connect$from, vertices$name)
to  <-  match( connect$to, vertices$name)
 
# Basic usual argument
ggraph(mygraph, layout = 'dendrogram', circular = TRUE) + 
  geom_node_point(aes(filter = leaf, x = x*1.05, y=y*1.05)) +
  geom_conn_bundle(data = get_con(from = from, to = to, values = connect$value), 
                   aes(colour = values),
                   alpha=0.2, 
                   # colour="skyblue", 
                   width=0.9) +
  scale_edge_color_gradient2(low = "red", high = "blue") +
  geom_node_text(aes(x = x*1.1, y=y*1.1, filter = leaf, label=name, angle = angle, hjust=hjust), size=1.5, alpha=1) +
  theme_void() +
  theme(
    legend.position="none",
    plot.margin=unit(c(0,0,0,0),"cm"),
  ) +
  expand_limits(x = c(-1.6, 1.6), y = c(-1.6, 1.6))


ggsave(paste0(model_wd, "/Plot_JSDM_mice_microbiome_bacteria_pa_20240214_spcorrelations.png"))


```

# 3. Species Associations plot - B

Plot the species associations categorizing the taxa by some grouping characteristic
We use a Hierarchical edge bundling plot

To create a bundling graph we need a hierarchy that is going to be used for the dendrogram. 
We add higher categories to our species, grouping them by env response, that we will use later for plotting

We also need to create the edges: data frame giving the hierarchical structure of the species,
grouped by species group


### Packages
```{r packages}
library(ggraph)
library(igraph)
library(dplyr)
library(grid)
library(png)
library(sf)
```


```{r load data}
# Taxa assoc
associations <- read.csv(paste0(model_wd, "/JSDM_mice_microbiome_bacteria_pa_20240214_sp_associations.csv"))
head(associations)

# groups
my_groups <- read.csv(**write file here**)

## create random grouping as example
# taxa
my_groups <- data.frame(species = unique(associations$species1))
# grouping
my_groups$group <- as.factor(rep(c("one", "two", "three"), times = 20))
# n sites it was detected
nsites <- data.frame(species = colnames(m$Y), 
                     nsites = colSums(m$Y))
my_groups <- my_groups %>% 
  left_join(nsites, by = "species")

my_groups

## Get sp names for the plots
associations$sp1 <- as.character(associations$species1)
# for (i in 1:length(birds.assoc$sp1)){
#   names1 <- strsplit(birds.assoc$sp1, '')[[i]][1]
#   names2 <- strsplit(birds.assoc$sp1, "_")[[i]][2] 
#   birds.assoc$sp1[i] <- paste(names1, names2, sep=".")
# }

associations$sp2 <- as.character(associations$species2)
# for (i in 1:length(birds.assoc$sp2)){
#   names1 <- strsplit(birds.assoc$sp2, '')[[i]][1]
#   names2 <- strsplit(birds.assoc$sp2, "_")[[i]][2] 
#   birds.assoc$sp2[i] <- paste(names1, names2, sep=".")
# }
  
head(associations, 20)  

# sp in traits
# Trdata$species <- as.character(Trdata$scientific)
# for (i in 1:length(Trdata$species)){
#   names1 <- strsplit(Trdata$species, '')[[i]][1]
#   names2 <- strsplit(Trdata$species, " ")[[i]][2] 
#   Trdata$species[i] <- paste(names1, names2, sep=".")
# }
# head(Trdata)

# Select association that are significant in the 95% CI
assoc_corr95 <- associations %>%
  filter(support < 0.025 | support > 0.975) %>%
  filter(species1 != species2) %>%
  dplyr::select(from = sp1, to = sp2, mean = mean, support = support) %>%
  mutate(correl = ifelse(mean < 0, "neg", "pos"))

nrow(assoc_corr95)
head(assoc_corr95)
```

```{r prepare hierarchical edge bundling plot}
# groups
d1 <- data.frame(from = "origin", to = levels(my_groups$group))
# d1 <- data.frame(from = "origin", to = levels(as.factor(my_groups$species)))

# Then we assign the species to the groups
d2 <- data.frame(from = my_groups$species, to = my_groups$species)
edges <- rbind(d1, d2)
nrow(edges)

## prepare nodes
#We use the number of sites that the species was detected for the size of the dots
vertices <- data.frame(name = unique(c(as.character(edges$from), as.character(edges$to))))
nrow(vertices)

vertices$size <- c(rep(NA, 4), my_groups$nsites)
vertices$group = edges$from[match( vertices$name, edges$to )]
head(vertices, 10)

# Reorder the data following the order we are interested in for representation:
# so all the species in the same group appear together
edges.ord <- edges[order(edges$from),]
vertices.ord <- vertices[order(vertices$group,vertices$name),]
head(vertices.ord, 10)

## calculate the ANGLE of the labels for plotting
vertices.ord$id=NA
myleaves <- which(is.na(match(vertices.ord$name, edges.ord$from))) # Select the rows with the final species, not the groups
nleaves <- length(myleaves) # This should be the number of species
vertices.ord$id[myleaves] <- seq(1:nleaves)

# First angle to calculate horizontal adjustment
vertices.ord$angle <- 90 - 360*vertices.ord$id/nleaves

# calculate the alignment of labels: right or left
# If I am on the left part of the plot, my labels have currently an angle < -90
vertices.ord$hjust<-ifelse( vertices.ord$angle < -90, 1, 0)

# flip angle BY to make them readable
vertices.ord$angle<-ifelse(vertices.ord$angle < -90, vertices.ord$angle+180, vertices.ord$angle)

head(vertices.ord, 10)
tail(vertices.ord, 10)
```

```{r create graph object and plot}
# give data graph format
mygraph <- graph_from_data_frame(d = edges.ord, vertices = vertices.ord)

# The connection object must refer to the ids of the leaves:
from <- match(assoc_corr95$from, vertices.ord$name)
to <- match(assoc_corr95$to, vertices.ord$name)


# we transform the correlations into positive or negative to better plot
connections <- assoc_corr95 %>%
  mutate (correlation = ifelse(mean > 0, "positive", "negative"))
connections

# load images
# bird1 <- readPNG(paste0(Images_wd, "/sparrow_darkgray.png"))
# g_bird1 <- rasterGrob(bird1, interpolate = TRUE)
# 
# bird2 <- readPNG(paste0(Images_wd, "/bird_silhouette1_darkgray.png"))
# g_bird2 <- rasterGrob(bird2, interpolate = TRUE)
# 
# urban <- readPNG(paste0(Images_wd, "/urban_gray.png"))
# g_urban <- rasterGrob(urban, interpolate = TRUE)
#  
# nature <- readPNG(paste0(Images_wd, "/opengreen_green.png"))
# g_nature <- rasterGrob(nature, interpolate = TRUE)
#  
# tree <- readPNG(paste0(Images_wd, "/trees_blue.png"))
# g_tree <- rasterGrob(tree, interpolate = TRUE)


# basic plot
my.plot <- ggraph(mygraph, layout = 'dendrogram', circular = TRUE) +
  geom_conn_bundle(data = get_con(from = from, to = to, values = connections$correlation), 
                   alpha=0.2, width=1.2, tension = 0.7, aes(colour=factor(values)), 
                   show.legend = FALSE) +
  geom_node_text(aes(x = x*1.2, y=y*1.2, filter = leaf, label = name, angle = angle, 
                     hjust = hjust, colour = group), size = 3.5, alpha = 0.7) +
  geom_node_point(aes(filter = leaf, x = x*1.07, y = y*1.07, colour = group, 
                      size = size, alpha=0.5)) +  
  expand_limits(x = c(-2, 2.5), y = c(-2, 2)) +
  # change edge color
  scale_edge_colour_manual(values = c("firebrick4", "midnightblue"), # there are only positive associations 
                           guide = guide_legend(override.aes = list(linetype = "solid", 
                                                                    colour = "grey50")),
                           name = "Association") +
  scale_edge_linetype("pointed") +
  # label color
  # scale_color_manual(values = c("grey20", "limegreen", "limegreen", "lightseagreen", "steelblue")) +
  # scale_color_manual(values = c("grey20", "darkgreen", "#FF6000")) +
    # scale_color_manual(values = c("#d8b365", "#f5f5f5", "#5ab4ac")) + 
  # label size
  scale_size_continuous(range = c(0.2,8),
                        breaks = c(10, 20, 50, 100, 200),
                        name = "Number of sites") +
  # labs(title="Species association network", caption="Source: simulated bird JSDM") +
  guides(size = guide_legend(order = 2), 
         # edge_colour = guide_legend (order = 1),
         colour = "none",
         alpha = "none") +
  # change theme  
  theme_minimal() +
  theme(legend.position = c(0.88, 0.2), 
        legend.background = element_rect(fill ="#F2F2F2", colour = "grey20"), 
        legend.title = element_text(size = 10), 
        legend.text = element_text(size = 8),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.title = element_blank(),
        axis.text = element_blank(), 
        plot.background=element_blank(), 
        panel.background = element_blank())

my.plot
# Add labels and pictures for diet categories
my.plot.labels <- my.plot +
  # add text
# my.plot +
  annotate(geom="text", x=1.6, y=1.7, label="Group 1: \nONE",
           color="grey20", fontface = "bold", size = 4) +
  # add box around text
  annotate("rect", xmin = 1.2, xmax = 2, ymin = 1.45, ymax = 1.9,
           color = "grey20", alpha = .2, fill = "grey20") +
  annotate(geom="text", x=-2.3, y=1, label="Group 2: \nTWO",
           color="#FF6000", fontface = "bold", size = 4) +
  annotate("rect", xmin = -2.8, xmax = -1.75, ymin = 0.8, ymax = 1.2,
           color = "#FF6000", alpha = .2, fill = "#FF6000") +
  annotate(geom="text", x=0, y=-2, label="Group 3: \nTHREEs",
           color="darkgreen", fontface = "bold", size = 4) +
  annotate("rect", xmin = -0.5, xmax = 0.5, ymin = -2.2, ymax = -1.8,
           color = "darkgreen", alpha = .2, fill = "darkgreen") 
  
# add images
# final.plot <- my.plot.labels +
#   annotation_custom(g_urban,
#                     xmin = 1.2, xmax = 1.55,
#                     ymin = 1.55, ymax = 2.5) +
#   annotation_custom(g_tree,
#                     xmin = -3, xmax = -1.5,
#                     ymin = 0.3, ymax = 0.7) +
#   annotation_custom(g_nature,
#                     xmin = -0.8, xmax = -1.5,
#                     ymin = -1.6, ymax = -2.3) +
#   annotation_custom(g_bird1,
#                     xmin = 2.8, xmax = 2,
#                     ymin = 0.2, ymax = 1) +
#   annotation_custom(g_bird2, 
#                     xmin = 1.5, xmax = 3,
#                     ymin = -1.2, ymax = -0.2) 
# 
#   final.plot

ggsave(paste0(model_wd, "/Plot_associations_bundle_groups_nsites.png"), 
       plot = my.plot.labels)
```


```{r}

library(ggraph)
library(tidygraph)
highschool

my_associations <- associations %>% 
  dplyr::select(from = sp1, to = sp2, mean, support)


my_associations <- my_associations %>% 
  filter(support > 0.95 | support < -0.95)

graph <- as_tbl_graph(my_associations)

# Not specifying the layout - defaults to "auto"
ggraph(graph) + 
  geom_edge_link(aes(colour = mean)) + 
  geom_node_point()



set.seed(1234)

nrow(associations)

d1 <- data.frame(from = "origin", to = paste("group", seq(1,10), sep=""))
d2 <- data.frame(from = rep(d1$to, each = 6), to = unique(associations$sp1))
hierarchy <- rbind(d1, d2)
 
# create a dataframe with connection between leaves (individuals)
all_leaves <- unique(associations$sp1)
connect <- data.frame(from = associations$sp1, to = associations$sp2, value = associations$mean)
  
# create a vertices data.frame. One line per object of our hierarchy
vertices  <-  data.frame(
  name = unique(c(as.character(hierarchy$from), as.character(hierarchy$to))) , 
  value = runif(71)
) 
# Let's add a column with the group of each name. It will be useful later to color points
vertices$group  <-  hierarchy$from[ match( vertices$name, hierarchy$to ) ]
 
 
# Create a graph object
mygraph <- graph_from_data_frame( hierarchy, vertices=vertices )
 
# The connection object must refer to the ids of the leaves:
from  <-  match( connect$from, vertices$name)
to  <-  match( connect$to, vertices$name)
 


# create a dataframe with connection between leaves (individuals)
all_leaves <- unique(associations$sp1)
connect <- data.frame(from = associations$sp1, to = associations$sp2, value = associations$mean)
 
# The connection object must refer to the ids of the leaves:
from <- match( connect$from, vertices$name)
to <- match( connect$to, vertices$name)

# plot
ggraph(mygraph, layout = 'dendrogram', circular = TRUE) + 
  geom_conn_bundle(data = get_con(from = from, to = to), alpha=0.2, colour="skyblue", tension = 0.9) + 
  geom_node_point(aes(filter = leaf, x = x*1.05, y=y*1.05)) +
  theme_void()
```

### Predicted responses 95CI by species

```{r}

```

### Caterpillar plots
```{r}
head(mpost$Beta)
mpost.beta.fox <- mpost$Beta[,1:14]
mpost.beta.raccoon <- mpost$Beta[,15:28]
mpost.beta.marten <- mpost$Beta[,29:42]

head(mpost.beta.raccoon)

colnames(mpost.beta.fox[[1]])
my_variables <- c("Intercept", "Sprin19", "Spring20", 
                  "No_compost", "Open_compost", "Garden_size", "Local_tree_cover", "Fence_height",
                  "Human_population", "Impervious_surface", "Noise", "Tree_cover", "D_CityBorder", "Cat")

#rename columns in betas
for (i in 1:length(mpost.beta.fox)){
  colnames(mpost.beta.fox[[i]]) <- my_variables
}
for (i in 1:length(mpost.beta.raccoon)){
  colnames(mpost.beta.raccoon[[i]]) <- my_variables
}
for (i in 1:length(mpost.beta.marten)){
  colnames(mpost.beta.marten[[i]]) <- my_variables
}

# for one model
mcmc_intervals(mpost.beta.fox) 

# for multiple models (ggplot)
head(mpost.beta.fox)

summary(mpost.beta.fox)$statistics
summary(mpost.beta.fox)$quantiles


# Put model estimates into temporary data.frames:
model1Frame <- data.frame(Variable = my_variables,
                          Coefficient = summary(mpost.beta.fox)$statistics[,1],
                          CI_low = summary(mpost.beta.fox)$quantiles[,1],
                          Q_25 = summary(mpost.beta.fox)$quantiles[, 2],
                          Q_50 = summary(mpost.beta.fox)$quantiles[,3],
                          Q_75 = summary(mpost.beta.fox)$quantiles[, 4],
                          CI_high = summary(mpost.beta.fox)$quantiles[,5],
                          Species = "Fox")

model2Frame <- data.frame(Variable = my_variables,
                          Coefficient = summary(mpost.beta.raccoon)$statistics[,1],
                          CI_low = summary(mpost.beta.raccoon)$quantiles[,1],
                          Q_25 = summary(mpost.beta.raccoon)$quantiles[, 2],
                          Q_50 = summary(mpost.beta.raccoon)$quantiles[,3],
                          Q_75 = summary(mpost.beta.raccoon)$quantiles[, 4],
                          CI_high = summary(mpost.beta.raccoon)$quantiles[,5],
                          Species = "Raccoon")

model3Frame <- data.frame(Variable = my_variables,
                          Coefficient = summary(mpost.beta.marten)$statistics[,1],
                          CI_low = summary(mpost.beta.marten)$quantiles[,1],
                          Q_25 = summary(mpost.beta.marten)$quantiles[, 2],
                          Q_50 = summary(mpost.beta.marten)$quantiles[,3],
                          Q_75 = summary(mpost.beta.marten)$quantiles[, 4],
                          CI_high = summary(mpost.beta.marten)$quantiles[,5],
                          Species = "Marten")


# Combine these data.frames
allModelFrame <- data.frame(rbind(model3Frame, model1Frame, model2Frame))  # We put marten first

allModelFrame$Variable <- factor(allModelFrame$Variable, levels = rev(my_variables))
allModelFrame$Species <- factor(allModelFrame$Species,levels = c("Raccoon", "Fox", "Marten"))
summary(allModelFrame)
# write.csv(allModelFrame, paste0(det_plots_wd, "/DET_fall_2springs_ModelFrame.csv"), row.names = FALSE)

allModelFrame[allModelFrame$Variable == "Human_population", ]

# Specify the width of your confidence intervals
# interval1 <- -qnorm((1-0.9)/2)  # 90% multiplier
# interval2 <- -qnorm((1-0.95)/2)  # 95% multiplier

# variables with CRI not overlapping 0
allModelFrame <- allModelFrame %>%
  mutate(significant = case_when(
    CI_low < 0 & CI_high < 0 ~ "Yes",
    CI_low > 0 & CI_high > 0 ~ "Yes", 
    TRUE ~ "No"))
 

# Plot
zp1 <- ggplot(allModelFrame, aes(group = Species, colour = Species)) + 
  geom_hline(yintercept = 0, colour = gray(1/2), lty = 2) + 
  geom_linerange(aes(x = Variable, ymin = CI_low,
                                ymax = CI_high, fill = significant),
                            lwd = 0.8, position = position_dodge(width = 1.2/2)) + 
  geom_linerange(aes(x = Variable, ymin = Q_25,
                                ymax = Q_75, fill = significant),
                            lwd = 1.5, position = position_dodge(width = 1.2/2)) + 
  geom_pointrange(aes(x = Variable, y = Coefficient, ymin = Q_25,
                                 ymax = Q_75, fill = significant),
                             lwd = 1/2, shape = 21, position = position_dodge(width = 1.2/2)) +
  scale_fill_manual(values = c("White", "black"))+
  
  scale_y_continuous(limits = c(-4.2, 3)) +
  coord_flip() + 
  scale_colour_viridis_d(option = "viridis", begin = 0.1, end = 0.9, 
                         guide = guide_legend(reverse = TRUE)) +
  theme(
    panel.background = element_rect(fill = NA),
    panel.grid.major = element_blank(), 
    axis.line = element_line(colour = "black")) +
    ggtitle("Urban mesocarnivores spatial responses, 95CI")


zp1

ggsave(plot = zp1, paste0(det_plots_wd, "/DET_coefs_fall_springs_allspecies.png"))

```

# 3. Species Associations - Circular plot 

Plot the species associations categorizing the taxa by some grouping characteristic
We use a Hierarchical edge bundling plot

To create a bundling graph we need a hierarchy that is going to be used for the dendrogram. 
We add higher categories to our species, grouping them by env response, that we will use later for plotting

We also need to create the edges: data frame giving the hierarchical structure of the species,
grouped by species group


### Packages
```{r packages}
library(ggraph)
library(igraph)
```

## Ploting

```{r load data}

# Taxa 
m <- readRDS(paste0(model_wd, "/JSDM_mice_microbiome_pa_20240311.rds"))
response <- as.data.frame(m$Y)

# Taxa assoc
associations <- read.csv(paste0(model_wd, "/JSDM_mice_microbiome_bacteria_pa_20240214_sp_associations.csv"))
head(associations)

# groups
my_groups <- read.csv(**write file here**)

## create random grouping as example
# taxa
my_groups <- data.frame(species = unique(associations$species1))
# grouping in three groups, as an example, no need 
my_groups$group <- as.factor(rep(c("one", "two", "three"), times = 20))
# n sites it was detected
nsites <- data.frame(species = colnames(response), 
                     nsites = colSums(response))
my_groups <- my_groups %>% 
  left_join(nsites, by = "species")

my_groups

## Get sp names for the plots
associations$sp1 <- as.character(associations$species1)
associations$sp2 <- as.character(associations$species2)
  
head(associations, 20)  

# Select association that are significant in the 95% CI
assoc_corr95 <- associations %>%
  filter(support < 0.025 | support > 0.975) %>%
  filter(species1 != species2) %>%
  dplyr::select(from = sp1, to = sp2, mean = mean, support = support) %>%
  mutate(correl = ifelse(mean < 0, "neg", "pos"))

nrow(assoc_corr95)
head(assoc_corr95)
```

```{r prepare hierarchical edge bundling plot}
# groups
d1 <- data.frame(from = "origin", to = levels(my_groups$group))

# Then we assign the species to the groups
d2 <- data.frame(from = my_groups$group, to = my_groups$species)
edges <- rbind(d1, d2)
nrow(edges)

## prepare nodes
#We use the number of sites that the species was detected for the size of the dots
vertices <- data.frame(name = unique(c(as.character(edges$from), as.character(edges$to))))
nrow(vertices)

vertices$size <- c(rep(NA, 4), my_groups$nsites)
vertices$group = edges$from[match( vertices$name, edges$to )]
head(vertices, 10)

# Reorder the data following the order we are interested in for representation:
# so all the species in the same group appear together
edges.ord <- edges[order(edges$from),]
vertices.ord <- vertices[order(vertices$group,vertices$name),]
head(vertices.ord, 10)

## calculate the ANGLE of the labels for plotting
vertices.ord$id=NA
myleaves <- which(is.na(match(vertices.ord$name, edges.ord$from))) # Select the rows with the final species, not the groups
nleaves <- length(myleaves) # This should be the number of species
vertices.ord$id[myleaves] <- seq(1:nleaves)

# First angle to calculate horizontal adjustment
vertices.ord$angle <- 90 - 360*vertices.ord$id/nleaves

# calculate the alignment of labels: right or left
# If I am on the left part of the plot, my labels have currently an angle < -90
vertices.ord$hjust<-ifelse( vertices.ord$angle < -90, 1, 0)

# flip angle BY to make them readable
vertices.ord$angle<-ifelse(vertices.ord$angle < -90, vertices.ord$angle+180, vertices.ord$angle)

head(vertices.ord, 10)
tail(vertices.ord, 10)
```

```{r create graph object and plot}
# give data graph format
mygraph <- graph_from_data_frame(d = edges.ord, vertices = vertices.ord)

# The connection object must refer to the ids of the leaves:
from <- match(assoc_corr95$from, vertices.ord$name)
to <- match(assoc_corr95$to, vertices.ord$name)

# we transform the correlations into positive or negative to better plot
connections <- assoc_corr95 %>%
  mutate (correlation = ifelse(mean > 0, "positive", "negative"))
connections

my.plot <- ggraph(mygraph, layout = 'dendrogram', circular = TRUE) +
  geom_conn_bundle(data = get_con(from = from, to = to, values = connections$correlation), 
                   alpha = 0.2, width = 1.2, tension = 0.7, aes(colour = factor(values)), 
                   show.legend = TRUE) +
  geom_node_text(aes(x = x*1.2, y = y*1.2, filter = leaf, label = name, angle = angle, 
                     hjust = hjust), size = 3.5, alpha = 0.7) +
  geom_node_point(aes(filter = leaf, x = x*1.07, y = y*1.07, #colour = group, 
                      size = size, alpha=0.5)) +  
  expand_limits(x = c(-2, 2.5), y = c(-2, 2)) +
  # change edge color
  scale_edge_colour_manual(values = c("firebrick4",
                                      "midnightblue"), # if there are only positive associations 
                            guide = guide_legend(override.aes = list(linetype = "solid",
                                                                     colour = "grey50"
                                                                     )),
                           name = "Association") +
  scale_edge_linetype("pointed") +
  scale_size_continuous(range = c(0.2,8),
                        breaks = c(10, 20, 50, 100, 200),
                        name = "Number of sites") +
  guides(size = guide_legend(order = 2),
         # colour = "none",
         alpha = "none") +
  # change theme  
  theme_minimal() +
  theme(legend.position.inside = c(0.88, 0.2), 
        legend.background = element_rect(fill ="#F2F2F2", colour = "grey20"), 
        legend.title = element_text(size = 10), 
        legend.text = element_text(size = 8),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.title = element_blank(),
        axis.text = element_blank(), 
        plot.background=element_blank(), 
        panel.background = element_blank())

my.plot

ggsave(paste0(model_wd, "/Plot_sp_associations_bundle_nsites.png"), 
       plot = my.plot)
```


